---
title: "Installing Rpath"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(eval = FALSE)
```

## Install/update R and needed R libraries
I tend to use Rstudio (GUI for R), you can use Rstudio or plain R Packages to Install in R (start R or Rstudio and enter the following) (you should only have to do this once in the class).
```{r packages}
install.packages("devtools")  # note: devtools is a big package, may take a while
install.packages("Rcpp")
install.packages("data.table")
install.packages("shiny")
```


## Install or update Rpath library 
From github using the github branch for the lab (will do this before each lab session to install new examples).  The 'ref' parameter is case-sensitive and must match the name of a branch in the github repo.
```{r rpath}
library(devtools)
install_github('NOAA-EDAB/Rpath', ref="lab_563_spring2022") 
```


## Possible mac compiler issue
On a mac, if installation dies with "ERROR: compilation failed for package ‘Rpath’" (or similar error) see: https://github.com/NOAA-EDAB/Rpath/issues/1


## Running the Rpath code test
```{r rpathtest}
  library(Rpath)
  library(shiny)
# Load an example Ecosystem 
# Find example ecosystem files (CSV format) for Gulf of Alaska (GOA)
  GOA_base <- system.file("extdata", "GOA_tiny_base.csv", package="Rpath", mustWork=T)
  GOA_diet <- system.file("extdata", "GOA_tiny_diet.csv", package="Rpath", mustWork=T)
  GOA_ped  <- system.file("extdata", "GOA_tiny_ped.csv",  package="Rpath", mustWork=T)

# Read in an unbalanced Ecopath model from csv file inputs
  path.unbal <- read.rpath.params(GOA_base, GOA_diet, GOA_ped)
  
# Make sure everything is installed and the files are loaded correctly 
# with this command
  check.rpath.params(path.unbal)
```
You should get the following output:  'Rpath parameter file is functional'


## Start Shiny app to explore ecosystem
```{r shiny}
shiny::runApp(system.file("labs/lab1", package="Rpath"), display.mode="normal")
```
This should open a browser window using your default browser(e.g. Chrome, Firefox, etc.) and the R terminal should read: 'Listening on http://127.0.0.1:' The browser window should display a drop-down list of functional groups, and display two bar graphs (mortality and diet) that change when the dropdown is changed. To exit browser mode, close the browser, or click back to the R terminal and hit ESC key.